### 2017-11-30 analysis/paper/paper.Rmd
clean up the reference list and citations

---
title: "A standard for the scholarly citation of archaeological data as an incentive to data sharing"
author:
  - name          : "Ben Marwick"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Anthropology, University of Washington, Seattle, 98115, WA, USA University of Wollongong, New South Wales, Australia"
    email         : "bmarwick@uw.edu"
  - name          : "Suzanne E. Pilaar Birch"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "University of Washington, Seattle, Washington, USA"
  - id            : "2"
    institution   : "University of Georgia,  Athens, Georgia,  USA"

date: "`r format(Sys.time(), '%d %B, %Y')`"

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
keywords          : ""
wordcount         : "" # '`r  prettyNum(unname(stringi::stri_stats_latex(readLines("paper.Rmd"))[4]), big.mark = ",")` (includes reference list )'

lang              : "english"
class             : "man"

output:
  datacitation::apa6_word

abstract: "" # |
   # "How do archaeologists share their research data, if at all? We review what data are, according to current influential definitions, and previous work on the benefits, costs and norms of data sharing in the sciences broadly. To understand data sharing in archaeology, we present the results of three pilot studies: requests for data by email; review of data availability in published articles, and analysis of archaeological datasets deposited in repositories. We find that archaeologists are often willing to share, but discipline-wide sharing is patchy and ad hoc. Legislation and mandates are effective at increasing data-sharing, but editorial policies at journals lack adequate enforcement. Although most of data available at repositories are licensed to enable flexible reuse, only a small proportion of the data are stored in structured formats for easy reuse. We present some suggestions for improving the state of date sharing in archaeology, among these is a standard for citing data sets to ensure that researchers making their data publicly available receive appropriate credit."
---

```{r, setup2, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  eval = TRUE,
  comment = "#>",
  fig.path = "../figures/"
)

library(datacitation)
library(tidyverse)

# my custom plot theme
theme_new <- function(base_size = 12){
  theme_bw(base_size = base_size) %+replace%
    theme(
      #axis.title.x=element_blank(),
      # axis.title.y=element_blank(),
      axis.title = element_text(size = 14),
      legend.key=element_rect(colour=NA, fill =NA),
      # panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, colour = "black", size=0.5),
      panel.background = element_rect(fill = "white", colour = "black"),
      strip.background = element_rect(fill = NA)
    )
}

```

Data are the building blocks of research, and access to data is necessary to fully understand and extend the work of others (Fienberg et al 1985). Yet archaeological data are often not available for a variety of reasons, including a perceived lack of credit for making data available. The goal of this paper is to present a standard for the scholarly citation of archaeological data, akin to the way we cite publications. The need for this has been motivated by Huggett’s (2017) findings of low levels of data citation and reuse in archaeology, and Clarke’s (2015) analysis that digital data preservation and management systems are under-developed in archaeology, despite our obligations as stewards of cultural heritage. A standard for data citation is important because it gives credit to the authors or compilers of the data, and so accords them visibility and recognition for their scholarly contribution in a way that is valued broadly in the discipline. A citation standard is also important for tracking the use of the data, and generating metrics about the impact of the dataset on the discipline. Understanding the use of resources such as datasets is important for allocating resources within archaeology and identifying novel research directions.

Archaeologists work with data in research, commercial, government, and other contexts. A full investigation of data in each of these contexts is beyond the scope of this paper, so we limit this study to what we are most familiar with, the research context. However, many of the practical details we discuss here will also be relevant to archaeologists working in other contexts. We begin by defining what we mean by ‘data’, and reviewing previous work on researchers’ attitudes to sharing data.  We motivate data sharing and citing behaviors by explaining why data sharing is beneficial to archaeology as a discipline, and to individual researchers. Then we survey some of the data sharing and citing norms that have emerged in other research areas. From these norms we derive a standard suitable for citation of archaeological data. Finally, we present some basic instructions on how to share archaeological data.

# What are Data?

Defining what data are is an ongoing research problem for information scientists (Leonelli 2016; Borgman 2012). Table \@ref(tab:summary-of-defs-of-data) summarizes some of the definitions debated in the philosophy of science literature. A more concrete starting point is this definition from the National Academies of Sciences:

"The term "data" as used in this document is meant to be broadly inclusive. In addition to digital manifestations of literature (including text, sound, still images, moving images, models, games, or simulations), it refers as well to forms of data and databases that generally require the assistance of computational machinery and software in order to be useful, such as various types of laboratory data including spectrographic, genomic sequencing, and electron microscopy data; observational data, such as remote sensing, geospatial, and socioeconomic data; and other forms of data either generated or compiled, by humans or machines." (Uhlir & Cohen, 2011, cited in Borgman 2015)

```{r summary-of-defs-of-data}
tbl <- readr::read_csv("../figures/tbl_definitions_of_data.csv")
knitr::kable(tbl, caption = "A selection of definitions of research data")
```

This is a useful definition because it emphasizes data as digital manifestations, and so suitable for transmission, storage and retrieval using the internet. Physical objects (such as samples, collections, hard copies of notes, etc.) are data also, but they are out of scope of this discussion because of the different challenges and requirements for access, archiving, and so on. The categories shown in Table \@ref(tab:types-of-data) are useful as a rough guide to what data can be, although this table is not comprehensive, and does not represent the full diversity of data used by archaeologists. Ultimately, data are whatever the community decides, as the The National Science Foundation (2010) notes in their commentary on Data Management Plans: "What constitutes such data will be determined by the community of interest through the process of peer review and program management. This may include, but is not limited to: data, publications, samples, physical collections, software and models." These definitions in policy are consistent with the relational account of data in philosophy of science scholarship, which holds that what data are is dependent on who uses it, how and for what purpose (Leonelli 2015; Daston and Galison 1992). How can these definitions and philosophical discussions inform our practice? They tell us that what counts as data is emergent from our professional interactions. Data are the materials we request when we do peer review of manuscripts and grant proposals, or when we make an informal request for more information by email. They are not exclusively objective archaeometric records, but include interpretive and derivative results. In short, data are anything another person needs to understand our work better.

To translate this into a more applied context, our practice of data sharing it to make available the computer files (e.g. containing text, raw structured data, images, etc.) that we analyze or otherwise use to create the figures, tables and quantitative results that we include in our reports and publications. These include our derived results and secondary data, such as taxonomic and typological classifications. This selection omits materials that could also be considered data according to the definitions noted above (e.g. field notes), but we feel our selection represents the most information-dense and meaningful items that are useful for our community of interest. These are also the items that have been most frequently requested from us by others in our professional interactions. Our citation standard is inclusive and imposes no constraints on what data can be, but we propose a few minimal qualities for how data should be organised that will help ground our discussion.

```{r types-of-data}
tbl <- readr::read_csv("../figures/tbl_types_of_data_nsb.csv")
knitr::kable(tbl, caption = "A selection of types of data encountered by archaeologists.")
```

# What are the Ethics of Data Sharing?

A full treatment of the ethical issues of data sharing is beyond the scope of this paper, but we can highlight some points specifically relevant to the question of when _not_ to share archaeological data. Much of the discussion on ethics and data sharing generally relates to appropriate attribution (assigning credit for the data) in data reuse (Duke et al. 2013). The default position for archaeological data, like data from many other sciences, is that they are part of the global human patrimony, and should be accessible to other scholars and the public (Zimmerman et al. 2003; Vitelli & Colwell-Chanthaphonh 2006). However, archaeologists, like other researchers working with local and indigenous communities, governments, and international collaborators, generate research data that is intellectual property with many stakeholders. With diverse stakeholders, archaeological data can become entangled in economics, nationalism, cultural politics, and identity (Nicholas and Wylie 2009). Often, one of these stakeholders is an indigenous community, and archaeologists may find differences between Western and indigenous conceptions of what data are, how they may be generated, used, shared, and 'owned' (Harding et al. 2012).

One of these differences is that Western researchers often have individual autonomy in sharing data, and once data are shared, it is free for all to use. In contrast, within indigenous communities, data are part of the group's identity and property, akin to tangible heritage (Nicholas 2014), with certain members entrusted to keep that data on behalf of the group, and sometimes it may be inappropriate for these people to share the data with other members of the group, or beyond the group (Tsosie 2007). Differences between Western and indigenous concepts or data are not only cultural - substantial inequalities exist in access to resources for converting archaeological data into economic and political benefit (e.g. graduate degrees, employment, publications, grants, promotions). On one hand, making archaeological data openly available benefits indigenous communities, because it removes some of the financial and technical barriers to accessing that data (e.g. subscription fees to journals). On the other hand, this may be little more than a useless courtesy (Hymes 1972), because simply making data openly available does not give indigenous people access to the contexts where the data can be used to generate economic and political benefits. Thus, open data has different values for Western and indigenous communities, and if archaeologists want to make a compelling case to their indigenous partners for data sharing, they should also make a commitment to reducing some of the inequalities in making use of the data, for example, by providing some training in data analysis, or adapting the data into a context that is valued by the indigenous community.

In short, there is more to the ethics of data sharing than avoiding risk of damage. Redacting the locations of archaeological sites to prevent looting is an obvious case of preparing data for sharing to reduce the risk of damage to the sites (and in some jurisdictions, mandated by law). However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive and exhaustive (Nicholas and Wylie 2012). It is crucial for archaeologists to discuss data sharing plans during negotiations with representatives authorized by the indigenous peoples whose cultural heritage is the subject of investigation. This negotiation process is important because only indigenous peoples themselves can identify potential adverse outcomes to data sharing, and they can do this only if they understand the proposed research and anticipated results. The concrete outcome of this process is often a Memorandum of Understanding between the indigenous community and the archaeologist's employer that documents the outcome of negotiations about data sharing. If the archaeologist's indigenous partners have reasons not to share data in public, then the archaeologist should respect that position, and note in publication that the data are restricted at the request of the indigenous collaborators.

# What are the Benefits of Data Sharing to the Individual Researcher?

While public and community benefits and costs are well-rehearsed and widely agreed on (Tables \@ref(tab:sharingcosts) and \@ref(tab:sharingbenefits)), they offer no, or only weak, tangible incentives to the individual researcher (Rowhani-Farid et al. 2017). Previous work has identified citation advantages as a tangible benefit to individual researchers who share data with publications, although the magnitude of the advantage varies greatly (see Table \@ref(tab:indivbenefits) for a summary). Increased publication productivity also appeared to be associated with data sharing behaviors. Across over 7,000 NSF and NIH awards, Pienta et al. (2010) reported that research projects with archived data produced a median of 10 publications, versus only five for projects without archived data.

```{r sharingcosts}
tbl <- readr::read_csv("../figures/tbl_reasons_not_to_share_data.csv")
tbl[,2] <- ""
names(tbl) <- c(names(tbl)[1], "")
knitr::kable(tbl, caption = "Summary of public and community costs of data sharing, from Stodden (2010) and Borgman (2012)")
```

.

```{r sharingbenefits}
tbl <- readr::read_csv("../figures/tbl_reasons_to_share_data.csv")
tbl[,2] <- ""
names(tbl) <- c(names(tbl)[1], "")
knitr::kable(tbl, caption = "Summary of public and community benefits of data sharing, from Stodden (2010) and Borgman (2012)")
```

.

```{r indivbenefits}
tbl <- readr::read_csv("../figures/tbl_evidence_of_sharing_benefits.csv")
knitr::kable(tbl, caption = "Summary of individual benefits to data sharing")
```

# What do Researchers Think about Sharing Data?

Sharing of data is a quality that is fundamental to the definition of science (cf. Merton’s (1942) communalism norm, where intellectual property rights are normally given up in exchange for recognition and esteem) and to data both as an abstract quality (cf. the definitions in Table \@ref(tab:summary-of-defs-of-data) that mention data as ‘portable’, ‘stable across different domains’) and as a concrete detail to be managed with policy. For example, for over a decade the National Science Foundation (2016) has had a data sharing policy that states "Investigators are expected to share with other researchers, at no more than incremental cost and within a reasonable time, the primary data, samples, physical collections and other supporting materials created or gathered in the course of work under NSF grants. Grantees are expected to encourage and facilitate such sharing." Surveys of researchers across different domains repeatedly show that people value the ideal of data sharing (Kratz and Strasser 2015). However, data sharing is not widespread in many disciplines, raising questions about why researchers do or do not share their data. Three recent studies have attempted to answer these questions, in the search for strategies to encourage sharing.

Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table \@ref(tab:sharingbenefits)). Stodden found that communitarian factors influence scientists to share their data, but private incentives determine when scientists choose not to share their code and data. In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process. Stodden’s (2009a, b) ‘Reproducible Research Standard’, summarized in Table \@ref(tab:stodden-rrs), is a convenient and robust copyright schema that is in wide use for many typical research projects.

```{r stodden-rrs, fig.cap="Summary of Stodden’s (2009b) 'Reproducible Research Standard'"}

tbl <- readr::read_csv("../figures/tbl_rrs.csv")
names(tbl) <- c("Contents", "Example", "License", "Comment")
knitr::kable(tbl, caption = "Summary of Stodden’s (2009b) 'Reproducible Research Standard'. This standard realigns legal rights with scientific norms by freeing scientific work from copying and reuse restrictions, with attribution. It encourages researchers to release their products with standardised instructions for reuse and attribution")
```

Other surveys have found similar results. Tenopir et al. (2011) received 1,329 responses to a survey conducted by the research team of the National Science Foundation-funded DataONE project. They found that when researchers did not share data it was due to reasons that include: insufficient time (54%), a lack of funding (40%), having no place to put the data (24%), a lack of standards (20%), and because their sponsor did not require it (17%). A remarkable observation by Tenopir et al. is that while 75% of respondents say they share data, only 36% of those respondents say their data are easy to access, while only 6% say they make all their data available. This suggests that most data sharing occurs in private directly between individual researchers, rather than in public via open repositories. A similar focus on private, peer-to-peer sharing is evident in interviews conducted by Wallis et al. (2013) with 43 researchers at the Center for Embedded Networked Sensing. They found that the most commonly cited conditions for sharing data were (1) first rights to publish results, (2) proper attribution to the data source, (3) familiarity between sharer and recipient, (4) funding agency expectations, and (5) the amount of effort required to share.

# What are the Current Norms of Data Sharing?

The surveys above show that data sharing practices tend to be a private activity, but where public sharing does occur, what forms does it take, and how do these vary across different fields of research?  Wouters and Haak (2017) surveyed 1,200 researchers about data sharing behaviors and found a spectrum from ‘intensive data-sharing fields’ to ‘restricted data-sharing fields’. Intensive data-sharing fields often depend on access to global databases, for example human genetics, while restricted data-sharing fields depend mostly on data collected by the individual researcher or their immediate collaborators, such as digital humanities. These differences in disciplinary norms have implications for data ownership (intensive: collectively owned, willing to share; restrictive: owned by the individual, reluctant to share) and the effort required to share (intensive: low effort because data are prepared for sharing as an integral part of the research process, restrictive: high effort because data are prepared for sharing after the research, or apart from it, in response to a request from another researcher). This intensive-restricted spectrum also maps on to Sawyer's (2008) analysis of data sharing in data-rich fields compared to data-poor fields. Data rich fields, such as astronomy, oceanography, and ecology, are have three common characteristics: pooling and sharing data resources is expected; the cumulative nature of work that builds on common data and common methods reduces the likelihood of alternative data sources and other methods being common; and the number of theoretical choices is limited because there is enough data to adequately test competing theories. Data-poor fields have three common characteristics: data are rare, or hard to get, or both; the types of data available often dictate the methods taken; and theories are relatively easy to generate and more difficult to validate. In data-poor fields, researchers often invest substantial time and resources on gathering their own data. The getting of data, and the quality of the data collected, is often a key determinant of the status of a researcher, leading to hoarding of data as an asset to be traded for professional benefit.

Archaeologists can be found all over this spectrum because of the diversity of archaeological research. However, our observations of data sharing in archaeology, and the nature of archaeological research generally, is that it is, for the most part, currently a restricted data-sharing and a data-poor field. There are few standardized and widely used methods for sharing and citing data, a challenge for archaeological data sharing that we hope this paper will help to address.

Even though there is a wide spectrum of data sharing behaviors, overall rates of data sharing by individual researchers are very low in many fields. Wicherts et al. (2006) emailed authors of 141 empirical articles recently published by the American Psychological Association (APA) journals to request a copy of the original data reported in the articles. They received 64 of the 249 datasets they requested, resulting in a data sharing rate of about 26%, despite all the authors of the articles in their study being bound to the APA's Ethical Principles, which state authors must ‘not withhold the data on which their conclusions are based from other competent professionals who seek to verify the substantive claims’ (American Psychological Association, 2001, p. 396). Valpaemel et al. (2015) similarly requested data from 394 papers, published in all issues of four APA journals in 2012 and found that 38% of authors responded with their data. They concluded that although the data sharing rate increased as compared to the study by Wicherts et al., their findings are ‘worrisome’ and indicate a 'poor availability of data'.

Savage and Vickers (2009) requested data from authors of ten papers published in PLOS journals, and received only one reply. Although this is a very small sample, it is notable because PLOS journals require authors to share data as a condition of publication: "PLOS journals require authors to make all data underlying the findings described in their manuscript fully available without restriction, with rare exception." (PLoS Medicine Editorial and Publishing Policies nd), and yet only one out of the ten authors actually complied with this requirement. Alsheikh-Ali et al. (2011) inspected 500 articles published in high-impact science journals in 2009 and found that only 47 papers (9%) deposited full primary raw data online. Similarly, Womak (2015) inspected 4,370 articles published in high-impact science journals in 2014 and found that 13% of articles with original data published make the data available to others.

The absence of a standardized and sustainable approach to data sharing is starkly demonstrated by Vines et al.’s (2014) study of data availability over time in biology publications. They examined the availability of data from 516 studies of plants or animal morphology between 2 and 22 years old. Vines et al. found that odds of a data set referred to in a publication being available fell by 17% per year. They found that the key factors responsible for the unavailability of data was that the data were lost or on inaccessible storage media.

# Do Archaeologists Share Data?

Here we present the results of three small pilot studies aimed at understanding how archaeologists share their data. First, we emailed a sample of authors of research publications to request access to the data behind their publication. Second, we surveyed a sample of journal articles to see what data was available. Third, we analyzed information about archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data). The small sample sizes of the first two of these pilot studies limit how generalisable the results are, but as the first of their kind for archaeology, they are valuable for proposing hypotheses for larger-scale investigations.

## Reproducibility and Open Source Materials for this Study

To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in Marwick (2017), we include the entire R code used for all the analysis and visualizations contained in this paper in our compendium at http://dx.doi.org/10.17605/OSF.IO/KSRUZ. Also in this version-controlled compendium are the raw data for all the tests reported here. All of the figures and quantitative results presented here can be independently reproduced with the code and data in this repository. In our compendium our code is released under the MIT license, our data as CC-0, and our figures as CC-BY, to enable maximum re-use (for more details, see Marwick 2017).

## Email Requests for Data in a Published Article

```{r import-email-data}
library(tidyverse)
email_survey <- readr::read_csv("../data/raw_data/email_survey.csv")

# overall rate of response
emails_out <- sum(!is.na(email_survey$`email sent date`))
emails_back <- sum(!is.na(email_survey$`response received date`))
email_response_rate <- round(emails_back / emails_out, 1) * 100
data_items_received <-  sum(!is.na(email_survey$`content recieved date`))
sharing_rate <-  round(data_items_received / emails_out, 1) * 100
```

Our sample focused on _Journal of Archaeological Science_ articles on experiments with stone artefacts that were published during `r min(email_survey$'publication year', na.rm = TRUE)` and `r max(email_survey$'publication year', na.rm = TRUE)`. We focused on stone artefact experiments because all research materials in these projects are generated by the researcher, so there are no other stakeholders to consider when determining how to share the data: the researcher has complete control over the data. There are also no risks of damage to people or property in sharing these data because they do not include archaeological sites or human subjects. Because we contacted authors of papers at the intersection of (1) containing eminently sharable data and (2) our _bona fide_ research interests (i.e. stone artefacts for BM), this is a targeted, non-random sample that may not be representative of the entire archaeological community. In contacting the authors, we requested the datasets for use in a graduate seminar with the aim of reproducing the published results and further exploring the data (this stage is still in progress). We promised to contact the authors if we discovered anything new, and not to share the data further without their permission. The data included with this paper is anonymized so the specific authors we contacted are not identifiable.

We sent out emails to the first authors of `r emails_out` stone artefact experiment papers, and received `r emails_back` replies, resulting in a `r email_response_rate`% response rate. We received `r data_items_received` responses that included data files, giving an overall sharing rate of `r sharing_rate`%. Our small sample size limits the robustness of our analysis, but points to some observations that may be worth following up in a larger study. We found no clear relationship of data sharing and date of publication. We do not have any insights into why authors of older papers decline to share, but three of the six responses for papers published in 2014-2015 said they declined to share data because they intend to use these data in a future publication. Other noteworthy reasons provided by respondents for declining to share include the author about to defend their thesis or about to go on vacation. We found no effect on the author's student status at time of publication on data sharing behaviors, with roughly equal proportions of students and non-students sharing and not sharing.

Similarly there are no clear patterns data sharing of among different journals. It is remarkable that the authors of the PLOS One paper we contacted declined to share because that refusal violates the policy of that journal at the time of publication of that paper (early 2014). Our finding is not unique - Chambers' (2017) examined 50 PLOS One papers on brain images published in November 2014 - May 2015 and found that only 38% of these papers archived their data in an open repository. Similarly, Nuijten et al. (2017) found that of 462 PLOS One articles that promised data availability, 29% of these did not have any available data. Unfulfilled promises to share data are not unique to PLOS One (Kidwell et al. 2016), but these studies highlight the need to resolve ambiguity among authors about what counts as data, and for editors and peer reviewers to more critically assess whether an author has made their data openly available.

We see similar rates of sharing in response to email requests among archaeologists to what have been described in other disciplines (Table \@ref(tab:sharing-rates), Tenopir et al. 2015; Vanpaemel et al. 2015). However, our result may overestimate the true rate because our sample did not include datasets that authors might not be able or willing to share because they contain sensitive location or cultural information. Analysis of the qualitative responses we received suggests two key reasons why authors are reluctant to share: they are fearful that they will lose the opportunity to produce further publications from those data (a typical concern of researchers in data-poor fields), and that their data are not in a form suitable for sharing and the task of organizing those data is not a high priority. These reasons not to share have also been documented in other research areas, noted above. The absence of patterns in the date of the paper, student status of the author and other variables (Figure \@ref(fig:email-figs)) may relate to more to our small sample size than a real absence of relationships in data sharing behavior.

```{r sharing-rates}
tbl <- readr::read_csv("../figures/tbl_studies_of_data_sharing.csv")
knitr::kable(tbl, caption = "Summary of a selection of previous studies of data sharing in various fields, from Spencer (2010). In the study by Peng et al., the authors did not attempt to contact other researchers, but noted the availability of data as described in the published article.")
```

```{r date-and-sharing, fig.cap="There is no obvious trend in data sharing becavious and date of publication"}
library(viridis)
library(scales) 
library(wesanderson)

years <- seq(min(email_survey$'publication year', na.rm = TRUE),
    max(email_survey$'publication year', na.rm = TRUE), 1)

sharing_by_year <-
email_survey %>%
  filter(!is.na(`email sent date`)) %>%
  group_by(`publication year`, `content recieved date`) %>%
  tally() %>%
  filter(!is.na(`publication year`)) %>%
  mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  ggplot(aes(`publication year`, n)) +
    geom_col(aes(fill = `content recieved date`),
             # need devtools::install_github("tidyverse/ggplot2") for this
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)], 
                    guide = FALSE) +
  scale_x_continuous(breaks = years,
                    labels = years) +
  xlab("Publication year") +
  scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
  theme_bw() +
  coord_fixed(ratio = 0.5)

```

```{r student, fig.cap="No relationship between student status and data sharing behavior"}
sharing_by_student <-
email_survey %>%
  filter(!is.na( `email sent date`)) %>%
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  group_by(
  `author is graduate student at time of publication?`,
  `content recieved date`) %>%
  tally() %>%
  filter(!is.na(`author is graduate student at time of publication?`)) %>%
  ggplot(aes(`author is graduate student at time of publication?`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
                    guide = FALSE) +
  scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
    xlab("Author was graduate student at time of publication?") +
    theme_bw() +
  coord_fixed(ratio = 0.05)
```

```{r author-country}
sharing_by_country <-
email_survey %>%
  filter(!is.na( `email sent date`)) %>%
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  group_by(
  `author home`,
  `content recieved date`) %>%
  tally() %>%
  filter(!is.na(`author home`)) %>%
  ggplot(aes(`author home`, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
                    guide = FALSE) +
    scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
    theme_bw() +
  xlab("Author location at time of publication") +
  coord_fixed(ratio = 0.25)
```

```{r journals, fig.cap="No relationship between sharing behavior and journal."}

email_survey <- 
  email_survey %>% 
  mutate(journal = ifelse(journal == "AMANT", 
                          "AQ",
                          journal))

sharing_by_journal <-
  email_survey %>%
  filter(!is.na( `email sent date`)) %>%
    mutate(`content recieved date` = if_else(!is.na(`content recieved date`),
  "yes", "no")) %>%
  group_by(
  journal,
  `content recieved date`) %>%
  tally() %>%
  filter(!is.na(journal)) %>%
  ggplot(aes(journal, n)) +
    geom_col(aes(fill = `content recieved date`),
             position = position_dodge(preserve = 'single')) +
  scale_fill_manual(values = wes_palette("Rushmore")[c(1,3)],
                    guide = guide_legend(title = "Data shared?")) +
    scale_y_continuous("Number of papers",
                     breaks= pretty_breaks()) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = c(0.8, 0.7)) +
  xlab("Journal name") +
  coord_fixed(ratio = 0.25)
```

```{r}
library(ggplot2)
library(gtable)
library(gridExtra)
library(grid)

library(cowplot)
email_sharing_plots <-
plot_grid(sharing_by_year,
                     sharing_by_student,
                     sharing_by_country,
                     sharing_by_journal,
          axis = "lt",
          ncol = 2,
          align = "hv",
          labels = "AUTO")

save_plot("../figures/email_sharing_plots.svg",
          email_sharing_plots,
          base_height = 8)
```

```{r email-figs, fig.cap="Summary of results of email requests for data. A. Responses by year of journal article publication, B. Student status of first author, C. Country location of first author, D. Journal (AAS = Archaeological and Anthropological Sciences, AQ = American Antiquity, ARCHMTRY = Archaeometry, JAS = Journal of Archaeological Science, JFA = Journal of Field Archaeology, JHE = Journal of Human Evolution, LT = Lithic Technology, SCI = Science."}
knitr::include_graphics("../figures/email_sharing_plots_hand_edit.png")
```

## Survey of Articles Published in the _Journal of Archaeological Science_

```{r}
jas_article_survey <- readr::read_csv("../data/raw_data/jas_article_survey.csv")

have_data_avail_statement <- nrow(jas_article_survey[jas_article_survey$`Data availability statement` != "none", ])
have_data_avail_statement_perc <- round(have_data_avail_statement / nrow(jas_article_survey) * 100,0)
data_are_avail <- nrow(jas_article_survey[jas_article_survey$`Can I access the data without contacting the authors?` %in% c("Yes", "yes"), ])
data_are_avail_perc <- round(data_are_avail / nrow(jas_article_survey) * 100,0)

data_in_table <-  nrow(jas_article_survey[grepl("Tab", jas_article_survey$`Where is the data?`), ])
data_in_table_perc <- round(data_in_table / data_are_avail * 100, 0)

data_in_supp <-  nrow(jas_article_survey[grepl("Sup", jas_article_survey$`Where is the data?`), ])
data_in_supp_perc <- round(data_in_supp / data_are_avail * 100, 0)

data_in_repo <-  nrow(jas_article_survey[grepl("www|dryad|osf", tolower(jas_article_survey$`Where is the data?`)), ])
data_in_repo_perc <- round(data_in_repo / data_are_avail * 100, 0)

data_type <- jas_article_survey %>%
  filter(`data type?` != "na") %>%
  group_by(`data type?`) %>%
  tally(sort = TRUE) %>%
  mutate(perc = round(n / sum(n) *100,1))

comp_n <- data_type %>%
  filter(`data type?`  == "compositional") %>%
  pull(n)

comp_perc <- data_type %>%
  filter(`data type?`  == "compositional") %>%
  mutate(perc = round(perc, 0)) %>%
  pull(perc)

dna_n <- data_type %>%
  filter(`data type?`  == "DNA sequence") %>%
  pull(n)

dna_perc <- data_type %>%
  filter(`data type?`  == "DNA sequence") %>%
  mutate(perc = round(perc, 0)) %>%
  pull(perc)

dates_n <- data_type %>%
  filter(grepl("dates", `data type?`)) %>%
  summarise(n = sum(n)) %>%
  pull(n)

dates_perc <- data_type %>%
  filter(grepl("dates", `data type?`)) %>%
  summarise(n = sum(n),
            perc = sum(perc)) %>%
  mutate(perc = round(perc, 0)) %>%
  pull(perc)

file_type <- jas_article_survey %>%
  filter(`What file format?` != "na") %>%
  group_by(`What file format?`) %>%
  tally(sort = TRUE) %>%
  mutate(perc = round(n / sum(n) *100,1))

xlsx_n <- file_type %>%
  filter(grepl("XLSX", `What file format?`)) %>%
  summarise(n = sum(n),
            perc = sum(perc)) %>%
  mutate(perc = round(perc, 0)) %>%
  pull(perc)

xlsx_perc <- file_type %>%
  filter(grepl("XLSX", `What file format?`)) %>%
  summarise(n = sum(n),
            perc = sum(perc)) %>%
  mutate(perc = round(perc, 0)) %>%
  pull(perc)

csv_n <- file_type %>%
  filter(grepl("CSV", `What file format?`)) %>%
  summarise(n = sum(n),
            perc = sum(perc)) %>%
  pull(n)

csv_perc <- file_type %>%
  filter(grepl("CSV", `What file format?`)) %>%
  summarise(n = sum(n),
            perc = sum(perc)) %>%
  mutate(perc = round(perc, 0)) %>%
  pull(perc)

doc_n <- file_type %>%
  filter(grepl("doc|DOC", `What file format?`)) %>%
  summarise(n = sum(n),
            perc = sum(perc)) %>%
  pull(n)

doc_perc <- file_type %>%
  filter(grepl("doc|DOC", `What file format?`)) %>%
  summarise(n = sum(n),
            perc = sum(perc)) %>%
  mutate(perc = round(perc, 0)) %>%
  pull(perc)

software <- jas_article_survey %>%
  filter(`Analysis software?` != "na") %>%
  group_by(`Analysis software?`) %>%
  tally(sort = TRUE) %>%
  mutate(perc = round(n / sum(n) *100,1))

script_n <- software %>%
  filter(grepl("^R$|.*R$|Python|MATLAB|OxCal", `Analysis software?`)) %>%
  summarise(n = sum(n),
            perc = sum(perc)) %>%
  pull(n)

script_perc <- software %>%
  filter(grepl("^R$|.*R$|Python|MATLAB|OxCal", `Analysis software?`)) %>%
  summarise(n = sum(n),
            perc = sum(perc)) %>%
  mutate(perc = round(perc, 0)) %>%
  pull(perc)

```

```{r}
# what is the total number of articles published in the population that we drew our sample from

# start http://doi.org.offcampus.lib.washington.edu/10.1016/j.jas.2016.11.003
# end   http://doi.org.offcampus.lib.washington.edu/10.1016/j.jas.2017.04.002

# start: Volume 78, February 2017, Pages 66-77
# end:   Volume 81, May 2017, Pages 90-100

# Vol, n_papers
# 81, 11
# 80,  9
# 79, 10
# 78, 18

# total: 48

```

In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` 48 articles published during February-May 2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviors. In this sample we found openly available raw data in `r data_are_avail` papers (`r data_are_avail_perc`%), even though only `r have_data_avail_statement` articles (`r have_data_avail_statement_perc`%) include a data availability statement. For papers where data was available, `r data_in_supp` papers (`r data_in_supp_perc`%) have data contained in supplementary files published with the paper, `r data_in_table` (`r data_in_table_perc`%) have all the raw data in tables in the text of the paper, and `r data_in_repo` (`r data_in_repo_perc`%) have data in an online repository. The most frequently shared type of data is compositional (e.g. element concentrations, n = `r comp_n`, `r comp_perc`%), followed by radiocarbon and luminescence age data (where the dates are the primary object of analysis, n = `r dates_n`, `r dates_perc`%) and DNA sequences (n = `r dna_n`, `r dna_perc`%). Generally, compositional data is presented in tables in the text, dating data is presented in supplementary files, and DNA data are in repositories (as is typical for DNA data) (Figure \@ref(fig:data-types-and-files)). Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, an open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%). When a scripting language such as R, Python, MATLAB or OxCal was used as the analysis software for a paper (n = `r script_n`, `r script_perc`% of the papers where would identify the software), only 2 papers also provided the script with the paper.

The _Journal of Archaeological Science_ has had a 'data disclosure' policy since at least 2013 (Editors, 2013) which states that "All data relating to the article must be made available in Supplementary files or deposited in external repositories and linked to within the article". The data availability rate of `r data_are_avail_perc`% reported in our pilot study reflects either weak enforcement of this policy or an incomplete understanding among editors, reviewers, and authors of how to interpret and implement this policy. The prominence of compositional and dating data among the shared data types suggests a least effort strategy by authors sharing data that do not require extensive cleaning and tidying (because the data are generated by instruments rather than entered by hand, such as artefact attribute data). The size of compositional and dating datasets means that they mostly fit easily into tables in the text or supplementary files, and do not require much effort to organize (e.g. very little metadata is required for others to use these data). The small proportion of papers in our sample using or sharing a scripting language suggests that unlike other fields (Eglen et al. 2017), archaeologists do not appear to recognize code for data analysis as a primary research object to share with the research community. Code written by the researcher for even the simplest operations should be shared along with data because code communicates decisions made about the data analysis that can help others understand, analyze, and reuse the data.

```{r}
# correlation of data type and location of sharing
file_type_location <-
jas_article_survey %>%
  mutate(file = strsplit(as.character(`What file format?`), ",")) %>%
  unnest(file) %>%
  filter(file != "na" & `Where is the data?` != "na") %>%
  select(file,  `Where is the data?`) %>%
  mutate(location = case_when(
    grepl("sup", tolower(`Where is the data?`)) ~ "Supplementary",
    grepl("tab", tolower(`Where is the data?`)) ~ "Table",
    grepl("www|dryad|osf", tolower(`Where is the data?`)) ~ "Repository",
  )) %>%
  mutate(file = case_when(
    grepl("DOCX|doc", file) ~ "DOCX",
    grepl("XLSX| XLSX", file) ~ "XLSX",
    TRUE ~ file
    ))

plot_file_type_location <-
file_type_location %>%
  group_by(file, location) %>%
  tally() %>%
  ggplot(aes(file, location)) +
  geom_point(aes(size = n),
             colour = wes_palette("Rushmore")[c(2)],
             shape = 16) +
  theme_bw() +
  xlab("") +
  ylab("") +
  scale_size_continuous(range=c(7,25)) +
  geom_text(aes(label = n)) +
  theme(legend.position = "none")
```

```{r}
# correlation of data type and location of sharing
data_type_location <-
jas_article_survey %>%
  filter(`data type?` != "na" & `Where is the data?` != "na") %>%
  select(`data type?`, `Where is the data?`) %>%
  mutate(data_type = case_when(
    grepl("comp", tolower(`data type?`)) ~ "Compositional",
    grepl("c14|osl", tolower(`data type?`)) ~ "Dating",
    grepl("dna", tolower(`data type?`)) ~ "DNA",
    TRUE ~ `data type?`
  )) %>%
   mutate(location = case_when(
    grepl("sup", tolower(`Where is the data?`)) ~ "Supplementary",
    grepl("tab", tolower(`Where is the data?`)) ~ "Table",
    grepl("www|dryad|osf", tolower(`Where is the data?`)) ~ "Repository"
  ))

plot_data_type_location <-
data_type_location %>%
  group_by(data_type, location) %>%
  tally() %>%
  filter(n >= 2) %>%
  ggplot(aes(data_type, location)) +
  geom_point(aes(size = n),
             colour = wes_palette("Rushmore")[c(2)],
             shape = 16) +
  theme_bw() +
  xlab("") +
  ylab("") +
  scale_size_continuous(range=c(7,25)) +
  geom_text(aes(label = n), colour = "black") +
  theme(legend.position = "none")
```

```{r data-types-and-files, fig.cap="Summary of relationships between file types, data types, and data locations. A. File types and data locations, B. Data types and data locations. "}
plot_data_type <- plot_grid(plot_file_type_location,
                            plot_data_type_location,
                      labels="AUTO", ncol = 1)
plot_data_type
```

## Datasets in Repositories Tracked by DataCite

```{r}
library(tidyverse)
# install.packages("rdatacite")

# osf project is at https://osf.io/ksruz/

# Original data aquisition on 14 Aug 2017
# library(rdatacite)
# dc_arch <- dc_search(q = "archaeology", rows = pretty(100000)[2])

# It doesn't get much of tDAR, so let's be more specific
# dc_arch_tdar <- dc_search(q = "10.6067", rows = pretty(100000)[2])
# library(osfr)
# upload_file("ksruz", "analysis/data/raw_data/dc_arch_tdar.rds", dest = "dc_arch_tdar.rds")

# Saved locally then uploaded to osf.io/tsm89 (too big for GitHub)
# saveRDS(dc_arch, "dc_arch.rds")
# saveRDS(dc_arch_tdar, "analysis/data/raw_data/dc_arch_tdar.rds")
# dc_arch <- readRDS("C:\\emacs\\dc_arch.rds")

# Download from osf.io with this:
# library(osfr)
# download_file("tsm89", private = TRUE)

dc_arch <- readRDS("../data/raw_data/dc_arch.rds")
dc_arch_tdar <- readRDS("../data/raw_data/dc_arch_tdar.rds")
dc_arch <- bind_rows(dc_arch, dc_arch_tdar)
```

```{r}
library(tidyverse)
library(stringr)
# glimpse(dc_arch)

dc_arch <-
dc_arch %>%
mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", rights)),
       the_text = gsub("(\\-)\\1+", "\\1", the_text),
       the_text = gsub("-$", "", the_text)) %>%
  mutate(cc = case_when(
    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
  )) %>%
  mutate(yyyy_date = as.numeric(substr(date, 1,4)),
         yyyy_created = as.numeric(substr(created, 1,4))) %>%
  mutate(yyyy_date = if_else(is.na(yyyy_date), as.numeric(publicationYear), yyyy_date)) %>%
  filter(yyyy_date > 1500) %>%
  filter(yyyy_date < 2018) %>%
  filter(!grepl("PANGAEA|Global Biodiversity", datacentre)) # pesky Mediterranean Data Archaeology and Rescue & Danish plants!
```

```{r}
# select cols of interest
xx <-
dc_arch %>%
  select( datacentre,
          allocator ,
          date,
          creator,
          description,
          language,
          contributor,
          rights,
          subject,
          format,
          publisher,
          yyyy_date,
          yyyy_created)

# tallies
datacentre <-  arrange(as_data_frame(table(xx[ , "datacentre"])), desc(n))
allocator  <-  arrange(as_data_frame(table(xx[ , "allocator"])), desc(n))
dates <-       arrange(as_data_frame(table(xx[ , "date"])), desc(n))
creator <-     arrange(as_data_frame(table(xx[ , "creator"])), desc(n))
language <-    arrange(as_data_frame(table(xx[ , "language"])), desc(n))
contributor <- arrange(as_data_frame(table(xx[ , "contributor"])), desc(n))
rights <-      arrange(as_data_frame(table(xx[ , "rights"])), desc(n))
publisher  <-  arrange(as_data_frame(table(xx[ , "publisher"])), desc(n))
file_format <-  arrange(as_data_frame(table(xx[ , "format"])), desc(n))
```

```{r first-dates}
dates <-
  dc_arch

# when was the first record created?
first_created <- min(dates$yyyy_created)
first_date <-  min(dates$yyyy_date, na.rm = TRUE)

# when is the break out
dates_tally <-
  dates %>%
  group_by(yyyy_date) %>%
  tally()

# when is that big peak, it's 2009
library(changepoint)
out <- cpt.meanvar(dates_tally$n, test.stat = "Poisson", method = "BinSeg")
breakouts <- dates_tally[cpts.ts(out), ]

# who is the top data source in each year?
top_data_sources <-
dates %>%
  group_by(datacentre) %>%
  tally(sort = TRUE) %>%
  slice(1:10) %>%
  pull(datacentre)

top_data_source_each_year <-
dates %>%
  group_by(datacentre, yyyy_date) %>%
  tally() %>%
  ungroup(datacentre) %>%
  group_by(yyyy_date) %>%
  slice(which.max(n)) %>%
  arrange(desc(yyyy_date))

 peak_year <-
   dates %>%
   group_by(yyyy_date) %>%
   tally(sort = TRUE) %>%
   slice(c(1,4)) %>%
   pull(yyyy_date)

 ymax <- 18000

date_plot <-
  dates  %>%
  filter(!is.na(yyyy_date)) %>%
  filter(datacentre %in% top_data_sources[1:5]) %>% 
  ggplot(aes(yyyy_date,
             fill = datacentre
             )) +
  geom_bar() +
  # guides(fill=FALSE) +
  scale_fill_manual(values = wes_palette("Rushmore")) +
  theme_new() +
  theme(legend.position = c(0.3, 0.5)) +
  scale_x_continuous(limits = c(1950,2017), expand = c(0,0)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite") +
  geom_vline(xintercept = peak_year, colour = "red") +
  annotate("text",
           x = peak_year - 1.5,
           y = ymax,
           label = peak_year,
           size = 3) # +
  # coord_fixed(ratio = 1/1000)

# change in each major repo over time
per_repo <-
dates  %>%
  filter(!is.na(yyyy_date)) %>%
  group_by(yyyy_date, datacentre)  %>%
  tally() %>%
  filter(n != 0,
         datacentre %in% top_data_sources[1:9]) %>%
    mutate(datacentre = str_wrap(datacentre, width = )) %>%
   ggplot(aes(yyyy_date,
              n
             )) +
  geom_line(size = 1)  +
  scale_fill_manual(values = wes_palette("Rushmore")) +
  scale_y_log10() +
  xlim(c(1990, 2017)) +
  xlab("Year DOI registered (1990-2017 only)") +
  ylab("Number of DOIs\nregistered per year") +
  theme_new() +
  facet_wrap(~ datacentre)

# show plots
datacite_temporal_plots <-
 plot_grid(date_plot,
           per_repo,
           labels="AUTO",
           ncol = 1,
           align = "hv",
           axis = "lr")

save_plot("../figures/datacite_temporal_plots.png",
          datacite_temporal_plots,
          base_height = 8)
```

```{r datacite-temporal-summary, fig.cap = "Change over time in archaeology datasets known to DataCite. A. Counts of datasets per year by the major data centres, red lines indicate major increases in the number of datasets in repositories. Note that the date on the horizontal axis indicates when the data was collected, not when it was deposited in the data centre, B. Temporal trends in the number of DOIs at individual data centres."}
knitr::include_graphics("../figures/datacite_temporal_plots.png")
```

```{r file-types}
# File types

# how many records with file type info?
file_types <- c("pdf|image|imagejpeg|jpg|tiff|png|gif|bmp|ai|illustrator|access|mdb|dbf|dbase|accdb|dbase|doc|rtf|word|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet|ply|archeolink|autocad|dwg")

number_of_records_with_file_type <-
file_format %>%
  filter(grepl(file_types, tolower(Var1))) %>%
  summarise(sum = sum(n)) %>%
  pull(sum)

perc_of_records_with_file_type <- round(number_of_records_with_file_type / nrow(dc_arch) * 100, 0)

# how many that are only PDFs?
have_a_pdf_in_there <-
file_format %>%
   filter(grepl("pdf", tolower(Var1))) %>%
   mutate(Var1 = tolower(gsub("application/|/", "", Var1)) )

perc_have_a_pdf <- round(sum(have_a_pdf_in_there$n) / number_of_records_with_file_type * 100, 0)

n_only_pdf <-  have_a_pdf_in_there %>%
    group_by(Var1) %>%
    tally(sort = TRUE) %>%
    filter(Var1 %in% c("pdf", "pdf,pdf", "pdfx,pdf", "pdfx")) %>%
    summarise(nnn = sum(nn)) %>%
    pull(nnn)

perc_only_pdf <-  round(n_only_pdf / number_of_records_with_file_type * 100, 0)

# how many that are only images?
have_an_image_in_there <-
file_format %>%
   filter(grepl("image|jpg|tiff|png|gif|bmp|ai|illustrator|imagejpeg", tolower(Var1))) %>%
   mutate(Var1 = tolower(gsub("application/|/", "", Var1)) )

n_only_image <-  have_an_image_in_there %>%
    group_by(Var1) %>%
    tally(sort = TRUE) %>%
    filter(Var1 %in% c("image", "jpg", "tiff", "png", "gif", "bmp", "ai", "illustrator", "imagejpeg", "imagetiff", "imagepng")) %>%
    summarise(nnn = sum(nn)) %>%
    pull(nnn)

perc_only_image <-  round(n_only_image / number_of_records_with_file_type * 100, 0)

# how many records contain a structured data file (csv, xls, dbf, etc)?
have_a_structured_data_file_in_there <-
  file_format %>%
  filter(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet", tolower(Var1))) %>%
   mutate(Var1 = tolower(gsub("application/|/", "", Var1)) )

n_structured_data_file <- have_a_structured_data_file_in_there %>%
  group_by(Var1) %>%
  tally(sort=TRUE) %>%
  summarise(nnn = sum(nn)) %>%
  pull(nnn)

perc_structured_data_file <- round(n_structured_data_file / number_of_records_with_file_type * 100, 0)

# count each file type once, even for records with multiple files types
file_format_single <-
file_format %>%
  filter(grepl(file_types, tolower(Var1))) %>%
  mutate(file_format = tolower(gsub("application|/", "", Var1))) %>%
  mutate(file_format_single = strsplit(as.character(file_format), ",|;")) %>%
  unnest(file_format_single)

# how many unique types?
how_many_unique_file_types <-
file_format_single %>%
  group_by(file_format_single) %>%
  tally(sort = TRUE)

file_format_simple <-
file_format_single %>%
  mutate(file_type = case_when(
    grepl("pdf", file_format_single) ~ "pdf",
    grepl("image|imagejpeg|jpg|tiff|png|gif|bmp|ai|illustrator", file_format_single) ~ "image",
    grepl("access|mdb|dbf|dbase|accdb|dbase", file_format_single) ~ "database",
    grepl("doc|rtf|word", file_format_single) ~ "doc",
    grepl("arcgis|shape|shp|mapinfo|mid|mif|surfer", file_format_single) ~ "GIS",
    grepl("xls|excel|csv|spreadsheet", file_format_single) ~ "spreadsheet",
    grepl("ply", file_format_single) ~ "3D",
    grepl("archeolink", file_format_single) ~ "archeolink",
    grepl("autocad|dwg", file_format_single) ~ "autocad"
  ))

file_format_simple_tally <-
file_format_simple %>%
  filter(!is.na(file_type)) %>%
  group_by(file_type) %>%
  tally(sort = TRUE) %>%
  mutate(perc = round(nn / sum (nn) * 100, 0))

file_format_combos_tally <-
file_format_simple %>%
  filter(!is.na(file_type)) %>%
  group_by(n) %>%
  filter(n_distinct(file_type) >= 2) %>%
  slice(1) %>% # look at the file_format col only
  group_by(file_format) %>%
  tally(sort = TRUE)

# which group has the highest *proportion* of structured data files?
structured_data_files <-
dc_arch %>%
    mutate(strutured_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet", tolower(format)), "has_structured_data", "no"))

proportion_structured_data_files <-
  structured_data_files %>%
  group_by(datacentre, strutured_data) %>%
  tally(sort = TRUE) %>%
  spread(strutured_data, n) %>%
  mutate(perc_structured = has_structured_data / (has_structured_data + no) * 100) %>%
  arrange(desc(perc_structured))

perc_struc_in_ads <-
  proportion_structured_data_files %>%
  filter(grepl("ADS", datacentre)) %>%
  pull(perc_structured) %>%
  round(., 2)

n_struc_in_ads <-
  proportion_structured_data_files %>%
  filter(grepl("ADS", datacentre)) %>%
  pull(has_structured_data)

perc_struc_in_dans <-
  proportion_structured_data_files %>%
  filter(grepl("DANS", datacentre)) %>%
  pull(perc_structured) %>%
  round(., 0)

n_struc_in_dans <-
  proportion_structured_data_files %>%
  filter(grepl("DANS", datacentre)) %>%
  pull(has_structured_data)

```

```{r}
library(tidyr)

structured_data_files_per_file <-
dc_arch %>%
    mutate(has_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet|pdf", tolower(format)), "has_data", "no")) %>%
  filter(has_data == "has_data") %>%
  mutate(format_simple = gsub("application|text|image|\\(|\\)| en ", "", format)) %>%
  separate(format_simple,
           sapply(1:50, function(i) paste0('f', i)),
           sep = "/|,|:|;|-|\\.| ") %>%
  select(datacentre, dataset_id, f1:f50) %>%
  gather(key = variable,
         value = value,
         -datacentre,
         -dataset_id,
         na.rm = TRUE) %>%
  select(-variable) %>%
  filter(value != "") %>%
  mutate(format = str_trim(tolower(value))) %>%
  filter(grepl("[a-z]+", format)) %>%
  select(-value) %>%
  mutate(format = gsub('ms', "", format)) %>%
  mutate(format = if_else(format %in% c("xls", "xlsx", "excel"),
                          "xls or xlsx",
                          format)) %>%
  filter(format != "") %>%
  distinct(datacentre, dataset_id, format)

# table(structured_data_files_per_file$datacentre)

top_file_formats <-
  structured_data_files_per_file %>%
  group_by(format) %>%
  tally(sort = TRUE) %>%
  slice(1:20) %>%
  filter(format != "p")

structured_data_files_per_file_tally <-
structured_data_files_per_file %>%
  group_by(format, datacentre) %>%
  tally(sort = TRUE) %>%
  filter(format %in% top_file_formats$format) %>%
  filter(datacentre %in% top_data_sources)

library(ggrepel)
plot_structured_data_files_per_file_tally <-
ggplot(structured_data_files_per_file_tally,
       aes(str_wrap(datacentre, 20),
           format)) +
      geom_point(aes(size = n),
             shape = 16,
             #alpha = 0.4,
             colour = wes_palette("Rushmore")[c(2)]) +
    geom_text_repel(aes(label = prettyNum(n, big.mark = ",")),
                  size = 2.5,
                  colour = "black",
                  segment.colour="black",
                  min.segment.length = unit(0.5, "lines"),
                  nudge_x = 0.3,
                  nudge_y = 0.5) +
  scale_size_continuous(range=c(2,25), guide = FALSE) +
    theme_bw() +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1)) +
  coord_flip() +
  xlab("Data centre") +
  ylab("File format")

ggsave("../figures/plot_structured_data_files_per_file_tally.png",
       plot_structured_data_files_per_file_tally,
       width = 10)
```

```{r, data-centre-file-types, fig.cap = "Relationships between data centres and file formats. The ten most frequently found file formats are shown here, in the data centres with these files from the top ten data centres by number of DOIs. Open file formats for structured data, such as CSV, are not common."}
knitr::include_graphics("../figures/plot_structured_data_files_per_file_tally.png")
```

```{r rights}
rights_summary <-
  rights %>%
  mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", Var1)),
         the_text = gsub("(\\-)\\1+", "\\1", the_text),
         the_text = gsub("-$", "", the_text)) %>%
  mutate(cc = case_when(
    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
  )) %>%
  select(cc, n) %>%
  filter(!is.na(cc)) %>%
  group_by(cc) %>%
  tally(sort = TRUE) %>%
  mutate(perc = round(nn / sum(nn) * 100, 0))

# what proportion is ADS or others?
rights_ads_perc <-
rights_summary %>%
  filter(cc == "ADS") %>%
  pull(perc)

rights_ads_n <-
rights_summary %>%
  filter(cc == "ADS") %>%
  pull(nn)

rights_eu_perc <-
rights_summary %>%
  filter(cc == "EU Open Access") %>%
  pull(perc)

rights_eu_n <-
rights_summary %>%
  filter(cc == "EU Open Access") %>%
  pull(nn)

```

Our third investigation of archaeologists' data sharing behaviors is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite. DataCite is the primary (but not the only) organisation that provides persistent identifiers, or DOIs, for research data. Although DataCite is an international organisation, our results are highly concentrated in just a few locations in Europe. The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY) (Figure \@ref(fig:datacite-summary)). Although the earliest data dates to `r first_date`, file deposition started and steadily increased during 1990-2008, as personal computers became more affordable and data easier to digitize. There is a substantial increase in 2009 when the DANS-EASY repository saw in influx of records from Dutch CRM firms. This can be explained as a result of Dutch archaeology legislation, the _Wet op de archeologische monumentenzorg_ (Wamz), in 2007 that formally obligated archaeologists in the Netherlands to archive their data via DANS within two years of completion of their projects (Keers et al. 2011). Another spike occurs in 2012 when tDAR had a large increase in deposits, likely due to them waiving their deposit fees for that year (Ellison 2012). After 2009 the volume of deposits declined in many repositories, with the exception of generic repositories such as Figshare, Zenodo and ResearchGate, which continued to grow (Figure \@ref(fig:datacite-temporal-summary)).

Of the `r prettyNum(sum(rights_summary$nn), big.mark = ",")` records with copyright information, the majority are released under the ADS conditions of use (n = `r prettyNum(rights_ads_n, big.mark = ",")`, `r rights_ads_perc`%), which is broadly similar to the CC-BY-NC (Moore and Richards 2015) (Figure \@ref(fig:data-licenses-summary)). The next most frequently used condition is the EU Open Access license (n = `r prettyNum(rights_eu_n, big.mark = ",")`, `r rights_eu_perc`%), similar to the CC-BY license. Creative Commons licenses such as CC-BY, CC-0 and others are used by less than 3% of records.

Many of the records are grey literature such as unpublished reports (Figure \@ref(fig:data-centre-file-types)), produced by commercial consultants or research projects. Of the `r prettyNum(number_of_records_with_file_type, big.mark = ",")` (`r perc_of_records_with_file_type`%) records that include information about the file types in the deposit, `r perc_only_pdf`% (n = `r prettyNum(n_only_pdf, big.mark = ",")`) contain only PDFs. Structured data files such as spreadsheets, databases or shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ",")`) records. The highest proportion of records with structured data files are found in the DANS-EASY repository (n = `r prettyNum(n_struc_in_dans, big.mark = ",")`, `r perc_struc_in_dans`%), followed by the ADS with `r perc_struc_in_ads`% (n = `r n_struc_in_ads`). Note that file format information is not present for all records in this sample, and some of the smaller repositories do not supply this information to DataCite at all, so it is possible that the true proportion of structured files in these records is higher than observed here.

From this pilot study of repository data we see that policy mandates in the UK and the Netherlands have been effective in populating repositories with archaeological data. With the 2012 spike in records deposited in tDAR we also see that archaeologists are sensitive to the price of sharing data. Nearly all records are made available under conditions that permit others to use the data while requiring that the original author of the data receive credit in reuse. In most cases this is because the repository hosting the data requires the author to choose one of these permissive licenses as a condition of making the deposit. However, although many items are _able_ to be reused because of the licenses, the ease of reuse of these items is low. Only a small fraction of the items deposited are machine-readable structured data files. The majority of items are PDFs, which require extensive manual handling to extract tabular data for convenient reuse and combination with other datasets.

```{r datacite-plots}
datacentre_plot <-
  datacentre %>%
  mutate(Var1 = str_wrap(Var1, 20)) %>%
  slice(1:10) %>%
ggplot(aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip() +
  xlab("Data centre") +
  ylab("") +
  theme_new() +
  theme(axis.text.y  = element_text(size = 8))

allocator_plot <-
ggplot(allocator[1:10,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip() +
  theme_new()

creator_plot <-
ggplot(creator[1:30,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip() +
  theme_new()

language_plot <-
  language %>%
  mutate(Var2 = ifelse(grepl("en|En", Var1), "en", Var1)) %>%
  mutate(Var2 = ifelse(grepl("de|ger", Var2), "de", Var2)) %>%
  group_by(Var2) %>%
  tally(sort = TRUE) %>%
  slice(1:5) %>%
ggplot(aes(reorder(Var2, nn),
           nn)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip()  +
  theme_new() +
  xlab("Language") +
  ylab("")

contributor_plot <-
ggplot(contributor[1:20,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip()  +
  theme_new()

rights_plot <-
ggplot(rights_summary,
       aes(reorder(cc, nn),
           nn)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip()  +
  theme_new() +
  xlab("License") +
  ylab("")

publisher_plot <-
ggplot(publisher[1:10,],
       aes(reorder(Var1, n),
           n)) +
  geom_col(fill = wes_palette("Rushmore")[4]) +
  coord_flip() +
  theme_new()

# show plots
# datacentre_plot
# allocator_plot
# creator_plot
# language_plot
# contributor_plot
# rights_plot
# publisher_plot

# Wow, we can nest cowplot::plot_grid in gridExtra::grid.arrange !! amazing and easy!
datacite_plots <-
arrangeGrob(datacentre_plot,
             plot_grid(language_plot,
                       rights_plot,
                       ncol=1,
                       align="v",
                       axis = "lr"),
             layout_matrix = rbind(c(1,2),
                                   c(1,2)),
             ncol = 2)

save_plot("../figures/datacite_plots.png",
          grid.draw(datacite_plots),
          base_height = 8)
```

```{r datacite-summary, fig.cap = "Distributions of data centers, languages (en = English, nl = Dutch, swe = Swedish, de = German, it = Italian) and licenses (see text for details) among archaeology datasets known to DataCite. Horizontal axis is the number of DOIs."}
knitr::include_graphics("../figures/datacite_plots.png")
```

```{r dates, eval = TRUE}
# why so many in 2009? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch

top_per_year <-
dates %>%
  group_by(yyyy_date, datacentre) %>%
   tally(sort = TRUE)  %>%
   top_n(n=1)

why_2013 <-
  xx %>%
  filter(as.numeric(substr(date, 1,4)) == 2013)  %>%
  group_by(datacentre) %>%
  tally(sort = TRUE)

why_2012 <-
  xx %>%
  filter(as.numeric(substr(date, 1,4)) == 2012)  %>%
  group_by(datacentre) %>%
  tally(sort = TRUE)

why_2011 <-
  xx %>%
  filter(as.numeric(substr(date, 1,4)) == 2011)  %>%
  group_by(datacentre) %>%
  tally(sort = TRUE)

why_2010 <-
xx %>%
  filter(as.numeric(substr(date, 1,4)) == 2010)  %>%
  group_by(datacentre) %>%
  tally(sort = TRUE)

why_2009 <-
  xx %>%
  filter(as.numeric(substr(date, 1,4)) == 2009)  %>%
  group_by(datacentre) %>%
  tally(sort = TRUE)

why_2008 <-
  xx %>%
  filter(as.numeric(substr(date, 1,4)) == 2008)  %>%
  group_by(datacentre) %>%
  tally(sort = TRUE)

# in 2009, among the Dutch, what is going on?
dutch_2009 <-
dc_arch %>%
  filter(yyyy_date == 2009,
         datacentre == "DELFT.EASY - DANS-EASY") %>%
  group_by(publisher) %>%
  tally(sort=TRUE)

# how about in ADS?
ads_2009 <-
dc_arch %>%
  filter(yyyy_date == 2009,
         grepl("ADS", datacentre)) %>%
  group_by(creator) %>%
  tally(sort=TRUE)

# compare to 2009
dutch_2008 <-
dc_arch %>%
  filter(yyyy_date == 2008,
         datacentre == "DELFT.EASY - DANS-EASY") %>%
  group_by(publisher) %>%
  tally(sort=TRUE)
```

```{r}

# look at some relationships
datacentre_rights_table <-
dc_arch %>%
  mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", rights)),
         the_text = gsub("(\\-)\\1+", "\\1", the_text),
         the_text = gsub("-$", "", the_text)) %>%
  mutate(cc = case_when(
    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
  )) %>%
  group_by(datacentre, cc) %>%
  filter(!is.na(cc))  %>%
  tally(sort = TRUE)

plot_datacentre_rights_table <-
datacentre_rights_table[1:11,] %>%
  ggplot(aes(cc, str_wrap(str_trunc(datacentre, side = "right", 35), 15))) +
  geom_point(aes(size = n),
             colour =  wes_palette("Rushmore")[1],
             shape = 16) +
  theme_bw() +
  xlab("License") +
  ylab("Data repository") +
  scale_size_continuous(range=c(7,25)) +
  geom_text(aes(label = prettyNum(n, big.mark = ","))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggsave("../figures/datacentre_rights_table.png",
       plot_datacentre_rights_table,
       height = 7)

# changes over time
# library(ggrepel)
#
# datacentre_by_year <-
# dc_arch %>%
#   group_by(yyyy_created, datacentre) %>%
#   tally(sort = TRUE) %>%
#   ungroup() %>%
#   filter(n > 150)
#
#
#   ggplot(datacentre_by_year,
#          aes(yyyy_created,
#              n,
#              colour = datacentre)) +
#   geom_line(size = 2) +
#   theme_bw()  +
#   scale_y_log10() +
#   theme(legend.position = "none")  +
#   facet_wrap(~ datacentre, scales = "free_y")

```

```{r data-licenses-summary, fig.cap = "Summary of licenses used by data centers with archaeological data"}
knitr::include_graphics("../figures/datacentre_rights_table.png")
```

```{r citation-events}
# get cross-ref events for dataset dois
# this is mostly citations of the datasets
library(crevents)

# contact the API and get the data...
# cred_query_safe <- safely(cred_query)
# events <-
# map(dc_arch$doi, ~cred_query_safe(rows = 10,
#                                   filter = list(`obj-id` = .x)))

# saveRDS(events, "../data/raw_data/cross_ref_events_2017_09_03.rds")
events <- read_rds("../data/raw_data/cross_ref_events.rds")

# Extract dataframes for analysis
events_df <-
map_df(events, ~.x$result$message$events)
# 366 obs 2 Oct
# 367 obs 3 Oct

# which provider is most cited?
# what sources do the most citing?

library(urltools)

events_df <- 
events_df %>% 
  mutate(doi_provider = str_extract(obj_id, 
                                    regex('(\\d+\\.\\d+)'))) %>% 
  mutate(doi_consumer = str_extract(subj_id, 
                                    regex('(\\d+\\.\\d+)'))) %>% 
  mutate(base_url_consumer = map_chr(subj.url, ~suffix_extract(domain(.x))[,1]))

# look at the top ones, manually look them up
top_doi_providers <- 
events_df %>% 
  group_by(doi_provider) %>% 
  tally(sort= TRUE) %>% 
  filter(n >= 10)

top_doi_consumers <- 
events_df %>% 
  group_by(doi_consumer) %>% 
  tally(sort= TRUE) %>% 
  filter(n >= 10)

doi_lookup <- 
tribble(  ~doi,       ~doi_name,
 "10.5281",   "Zenodo",
 "10.5284",   "Medieval Archaeology",
 "10.6067",   "tDAR",
 "10.15468",  "GBIF",
 "10.14279",  "Technische Universität Berlin",
 "10.17026",  "EASY-DANS",
 "10.11141",  "Internet Archaeology",
 "10.1080",   "Various journal publishers") # OUP, CUP, Sage, Wiley, Springer

# join names to main table
events_df_names <- 
  events_df %>% 
  left_join(doi_lookup, by = c("doi_provider" = "doi")) %>% 
  rename(doi_provider_name = doi_name) %>% 
  left_join(doi_lookup, by = c("doi_consumer" = "doi"))  %>% 
  rename(doi_consumer_name = doi_name) %>% 
  mutate(base_url_consumer = suffix_extract(base_url_consumer)$domain) %>% 
  mutate(doi_consumer_name = if_else(is.na(doi_consumer_name), 
                                     base_url_consumer, 
                                     doi_consumer_name))

# cross-tab
events_df_names_ct <- 
events_df_names %>% 
  group_by(doi_provider_name, 
           doi_consumer_name) %>% 
  tally() %>% 
  filter(!is.na(doi_provider_name)) %>% 
  filter(!is.na(doi_consumer_name)) %>% 
  spread(doi_consumer_name, n, fill = 0)

library(wesanderson)
plot_citation_table <-
events_df_names %>% 
  group_by(doi_provider_name, 
           doi_consumer_name) %>% 
  tally() %>% 
  filter(!is.na(doi_provider_name)) %>% 
  filter(!is.na(doi_consumer_name)) %>% 
  ggplot(
         aes(str_wrap(str_trunc(doi_provider_name, side = "right", 35), 15),
             str_wrap(str_trunc(doi_consumer_name, side = "right", 35), 15))) +
  geom_point(aes(size = n),
             colour =  wes_palette("Rushmore")[1],
             shape = 16) +
  theme_bw() +
  xlab("Data source") +
  ylab("Cited by") +
  scale_size_continuous(range=c(5,25)) +
  geom_text(aes(label = prettyNum(n, big.mark = ","))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggsave("../figures/doi_citation_table.png",
       plot_citation_table,
       height = 7)
```

```{r citation-time-series}
citation_time_series_plot <-  
events_df_names %>% 
  select(occurred_at) %>% 
  mutate(occurred_at = as.Date(occurred_at)) %>% 
ggplot(aes(occurred_at)) +
  geom_histogram() +
  scale_x_date(date_breaks = "6 month") +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  xlab("Date of citation event") +
  ylab("Number of citations\nper day")

ggsave("../figures/citation_time_series_plot.png",
       citation_time_series_plot,
       height = 3)

```

```{r combine-citation-plots}
# combine citation plots
library(cowplot)

combined_citation_plots <- 
ggdraw() +
  draw_plot(plot_citation_table) +
  draw_plot(citation_time_series_plot,
            x = 0.25, y = -0.1, scale = 0.4)

ggsave("../figures/combined_citation_plots.png",
       combined_citation_plots,
       height = 7, width = 10)
```

We can get some insights into the impact of repository data by querying the Crossref (2017) Events Data service, a freely accessible DOI citation dataset produced by periodic scanning of a variety of online sources. Crossref is an official DOI registration agency, focused on issuing DOIs for scholarly materials such as journal articles, books, ebooks, etc. Another source for data citation data is Thomson Reuters' Data Citation Index (Pavlech 2016), but this resource is not free, and we did not have access to a subscription. Of the `r prettyNum(length(dc_arch$doi), big.mark = ",")` DOIs in our sample, we found a total of `r nrow(events_df)` citations of `r length(unique(events_df$obj_id))` unique DOIs in the Crossref events data. The majority of these citations were found on Wikipedia (Figure \@ref(fig:citation-plot)), followed by other data repositories (i.e. cross-citations from one data set to another in the same repository). The largest source of citations is the Zenodo repository, followed by articles in the journal _Medieval Archaeology_, which appear here because volumes 1-50 are hosted by the Archaeology Data Service. According to the Crossref data, citation of archaeological data in repositories is a recent phenomena, with the majority in our sample occurring over the last twelve months. We found only one example of a dataset DOI cited in a journal article (McCoy and Codlin 2016). This may be an underestimate, given Huggett's (2017) finding that 56 of 476 archaeology data sources in Thomson Reuters' Data Citation Index have citations in Web of Science databases. However, Huggett notes that the Thomson Reuters citation data is unreliable, for example, with incomplete citations, and citations that pre-date the cited dataset. Huggett also shows an example of archaeologists citing datasets but omitting the DOI (e.g. Nugent and  Williams 2012), so the citation is not captured by services such as Crossref or Thomson Reuters. These malformed citations add to the difficulty of accurately measuring data citation in archaeology.

```{r citation-plot, fig.cap = "Citation of archaeological data DOIs tracked by the Crossref Events Data service. Inset shows the distribution of these citations over time."}
knitr::include_graphics("../figures/combined_citation_plots.png")
```

# How to Share Archaeological Data?

In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011). List of trustworthy repositories are maintained by the editors of the journal _Scientific Data_ (n.d) and <http://www.re3data.org>, a global registry of research data repositories (Pampel et al. 2013). University librarians are another resource for advice on choosing repositories and preparing data for sharing. There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al. 2012; Wilkinson et al. 2016). Table \@ref(tab:desirable) summarizes some of these, and recommendations specifically written for archaeologists can be found in Archaeology Data Service / Digital Antiquity (2011) and ARCHES project. (n.d.). Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data. However, we should not let the perfect be the enemy of the good. In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations. While we recognize that our data sharing practices fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.

Data sharing is a relatively new behavior for most archaeologists, so it is important to clarify what does not count as data sharing. Making data available on a web site that is not a trustworthy, persistent-identifier-issuing repository is not data sharing. For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any time, the data can change in untracked ways, and published links to the files are broken when these websites are reorganized, making the files inaccessible. Sharing data via supplementary files attached to a journal article is problematic because they are inaccessible to non-subscribers for paywalled journals, because publishers often alter the data file formats and file names during the production process, and because publishers own that data, and the original author may need to request permission to reuse their own data. For example, machine-readable formats such as CSV are sometimes presented as PDF or Microsoft Word files in supplementary files. The modification of supplementary file names during the article production process can break scripts that attempt to read files by referring to file names as they were prior to the publisher handling the files.

Finally, authors who write in their publications that 'the data are available by request' are not sharing data. Not only are the data not actually available, but there is no way to enforce that the authors will actually honor any requests for their data. A simple strategy for avoiding these pseudo-sharing behaviors is to deposit data files in a trustworthy data repository, and use persistent links to the files in reports and manuscripts. In our own work we have used figshare.com, zenodo.org, osf.io, neotomadb.org, opencontext.org and our University repositories. However, there are many other options, and a full discussion of how to choose the ideal data repository is beyond the scope of this paper. We recommend seeking advice from librarians and funding agencies to help with finding the most suitable service.

# How to Cite Archaeological Data?

The Joint Declaration of Data Citation Principles is the current standard in many disciplines for data citation within any research object (Data Citation Synthesis Group 2014). The result of these principles is a citation formation as follows (the exact order of the items may vary slightly across different referencing styles):

Creator (PublicationYear). Title. Version. Publisher. Identifier

The 'PublicationYear', 'Title' and 'Publisher' are familiar from traditional citation formats, except that the publisher may not be a traditional book or journal publisher, but a dedicated data repository organisation. 'Creator' is equivalent to the author or primary collector and arranger of the data (e.g. Open Context or tDAR). The 'Version' and 'Identifier' fields are unique to data citation, and can take a variety of forms. The identifier is a special type of persistent internet address. The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher, but not necessarily in a human-readable format), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes. There are several varieties of these in use, with DOIs (digital object identifiers) currently the most common form. When a data publisher changes the internet address of a data set, they are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.

A DOI consists of three parts:

- the proxy, usually https://doi.org but sometimes also https://hdl.handle.net/ or just doi:
- the prefix, usually a string of numbers that is unique to the organisation publishing the DOI. For example, tDAR prefix is 10.6067
- the suffix, a random-looking string of any combination of characters and numbers that uniquely identified the file or bundle of files, this is also known as a local identifier (local ID) or an 'accession', an identifier which, by itself, is only guaranteed to be locally unique within the database or source (McMurry et al. 2017)

A DOI alone is not sufficient for citing a dataset because, like many things on the internet, the files can change over time. For example we might upload data files to the repository when we submit a paper for peer review, then get a DOI and cite that in our manuscript so the peer reviewers can see them (some services provide a private URL that can be shared with reviewers but is not publicly accessible or discoverable by a search engine). Then when we resubmit our manuscript after peer review, we might update the repository to include additional data files relating to additional analyzes requested by the reviewers. We still have the same DOI in our manuscript for this data set, but the files have changed because of the revisions we did in response to the peer reviewers' comments. Most data repositories will keep each version of the files that you upload, so in this scenario there will be at least two versions of the files in our data repository. Even if we do not follow this exact sequence of updating the files, we may still change them in the future, after our paper has been published. Because of this potential to change the files stored at the DOI, it is necessary to identify specific file versions in our citation of the data. A simple method to record the version is to include the date that the data files were accessed, as when citing any other resource found on the internet. Some repositories also provide DOIs that have a version number suffix like '.v2', '.2', '/2'. However this is not formally a functionality of the DOI system, so including the date accessed is a more robust and general method of indicating the version. Technological solutions to issues of transience, permanence and granularity in data citation are still in development.

When citing data in the text of a manuscript or report, the ideal practice is to cite the dataset and the traditional publication that first described it, like so (Penev et al. 2017):

"This paper uses data from the [name] data repository at http://doi.org/xxx/xxx (Author YYYY), first described in Author YYYY"

For example:

"This paper uses data from the Open Science Framework data repository at http://doi.org/10.17605/OSF.IO/32A87 (Marwick 2017), first described in Marwick et al. 2017."

Citing both the dataset and the publication helps to resolve ambiguity about the source of the data, and gives context to the data. Furthermore, the authorship of the data files may differ from those of the author's paper describing the data, and it is important that both authors (or sets of authors) receive credit for their work through the normal mechanisms of citation recognition. In cases where an author has published their data, and then makes an intellectual contribution to a separate study lead by others that uses the published data, then the data author should also be a co-author of the separate study. If other researchers use published data in their publication without any input from the data author, then citation of the data is all that is required, and adding the data author as a co-author is not required.

In the paper in which these data are first described, the citation statement should be included in the body of the paper, in the Material and Methods section. In addition, the formal data reference should be included in the paper's reference list, using the recommended journal's reference format. For example, here is a statement that would appear in the paper in which the data are first presented:

"The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://doi.org/10.17605/OSF.IO/32A87 (Marwick 2017)."

With the reference list of that paper including Marwick (2017) as:

Marwick, B. (2017). Dataset and R code for "Movement of lithics by trampling: An experiment in the Madjedbebe sediments", http://doi.org/10.17605/OSF.IO/32A87, Open Science Framework, Accessed 7 Sept 2017.

# Summary and Conclusion

The status of research data is slowly changing in many fields from a gift-culture where data are privately-held commodities traded for co-authorship, invitations, and other status tokens (Hagstrom 1982; Wallis et al. 2013), to part of the public commons, accessible to anyone with an internet connection. This change has been fastest in data-rich 'big science' fields, such as astronomy, physics, and oceanography, that use expensive instruments that generate data shared with large teams of collaborators to produce new discoveries (Reichman and Uhlir 2003). However, in 'smaller' data-poor sciences such as archaeology, where collaborative groups are small, data sharing is still mostly ad hoc and dis-aggregated. Our pilot studies, while small and limited in many ways, support this statement with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations. We found that legislation requiring data sharing resulted in a substantial increase in the rate of data deposition, and that archaeologists are sensitive to repository fees. In examining archaeological data deposited in repositories, we found that most of the contents are PDF files of reports, and only a very small fraction of the data files are structured data such as spreadsheets. Archaeologists still have a long way to go in making their data easier to reuse by sharing data in plain text structured data formats such as CSV. We found that citation of datasets, although problematic to measure, appears to be almost non-existent in archaeological literature. Table \@ref(tab:desirable) summarizes nine key tasks archaeologists should do to their data to optimize its discovery and reuse.

```{r desirable}
tbl <- readr::read_csv("../figures/tbl_rules_of_data_sharing.csv")
knitr::kable(tbl, caption = "Nine key things to do before sharing your data, from Hart et al. (2016) and White et al. (2013)")
```

Among the many challenges of increasing the rate of data sharing in archaeology and other 'small' sciences is that data sharing is unfunded, unrewarded and is only rarely required. Who is responsible for changing this? There is no simple answer, but we believe the task must be shared between individuals and institutions, with individuals pushing for institutional change. The challenge of funding can be tackled by practical action from members of professional societies. For example, the Society of American Archaeology should communicate to funding bodies, such as the National Science Foundation, about the importance of allocating funds specifically for data sharing, and providing reliable information on how to choose and and use data repositories. The challenge of requiring data sharing can also be addressed by funding bodies and federal agencies, through more rigorous evaluation and enforcement of data management plans during the review process. However, changes in the policies of professional societies and other agencies are primarily responses to changes in norms in the community of researchers. In practice, individual researchers are responsible for communicating the value of data sharing to professional societies and funding agencies (e.g. by holding peers to high standards of data sharing during grant reviews). Changes in individual practice will demonstrate a change in norms, and motivate our institutions to formalize these new data sharing values through policies and regulations.

One area where changes in individual practice may be especially effective is among journal editors and editorial boards. For example, editors of archaeology journals should adopt the Transparency and Openness Protocols of the Center for Open Science (Nosek et al. 2015). Editors should then educate their editorial boards members, peer reviewers, and contributors about their obligations to these protocols. Editors should only accept papers that, minimally, provide the raw data used to produce each table and visualisation in the manuscript. Changes in journal-specific submission guidelines can result in substantial changes in authors’ research practices (Giofrè et al 2017). For example, if archaeology journals adopt incentives such as badging (Kidwell et al. 2016), and require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analyzed in the article (or if the data cannot be shared, an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published.

Individual researchers can also contribute to changing the norms on data sharing via several practical actions suitable for typical day-to-day work:

- When preparing research for publication, upload data to a trustworthy repository and cite the data in the manuscript.      
- When writing peer reviews of journal articles, we can tell editors that we require data to be available for papers that we read, and that the manuscripts we review must include citations of datasets analyzed in the manuscript. We can write in our peer reviews that data need to be available for us to make a proper assessment of the manuscript.    
- When reviewing grant applications, we can critically evaluate data management plans, and report on the applicants' past record on data sharing. We can offer helpful suggestions about how to improve data availability.     
- When teaching students, we can teach them how to prepare data for sharing, and how to find and use publicly available datasets to give students experience analyzing and writing about shared data, and to cultivate the expectation that data sharing is a normal part of archaeological research.     
- When evaluating candidates for hiring and promotion, ensure that commitments to data sharing are considered.

The challenge of data sharing as an unrewarded behavior was the primary motivation for this article. By specifying a standard for the scholarly citation of archaeological data we provide a method for giving credit to the authors or compilers of the data. This rewards those authors with visibility and recognition for their effort with citations, a currency that is valued broadly in the discipline. The data citation standard outlined above is easy for humans and search engines to understand, and is highly functional, with a full internet address to the data contained in the citation. The standard can accommodate the addition of additional information required by different citation styles and publishers' requirements. The data citation standard is not dependent on any specific choices of software, archive, data producer, publisher, or author and is easy to adapt into existing writing and research workflows. There are still many unresolved challenges in data sharing, such as how to manage high volumes of data generated by photogrammetry. However, we are confident that solutions to many of these challenges will emerge as data sharing is increasingly valued by the research community.

# Acknowledgements

This paper emerged out of discussions at the 2017 SAA meeting, specifically the forum ‘Beyond Data Management: A Conversation about “Digital Data Realities”’, organised by Sarah Kansa and Eric Kansa, and the SAA publications committee meeting. Thanks to Canan Cakirlar for insights into the history of archaeological data sharing in the Netherlands, and Stefano Costa for many helpful suggestions. BM was supported by an Australian Research Council Future Fellowship (FT140100101). Thanks to Stefanie Crabtree and the other four peer reviewers for their constructive suggestions.

# References Cited

Alsheikh-Ali A.A., Qureshi W., Al-Mallah M.H. and J. Ioannidis
2011. Public Availability of Published Research Data in High-Impact Journals. PLOS ONE 6(9): e24357. DOI:10.1371/journal.pone.0024357, accessed 7 September 2017.
American Psychological Association
2001.	Ethical Standards for the Reporting and Publishing of Scientiﬁc Information. In Publication Manual edited by American Psychological Association, pp. 387–396. Washington, DC
ARCHES project
n.d. The Standard and Guide to Best Practice in Archaeological Archiving in Europe. Electronic document, http://archaeologydataservice.ac.uk/arches/Wiki.jsp?page=ARCHES%20Introduction, accessed on 7 September 2017.
Bogen J, and J. Woodward
1998.	Saving the Phenomena. The Philosophical Review 97(3): 303–352
Borgman, C. L. 
2015.	Big Data, Little Data, No Data: Scholarship in the Networked World. MIT Press, Massachusetts. 
Borgman, C. L. 
2012.	The Conundrum of Sharing Research Data. Journal of the American Society for Information Science and Technology, 63(6), 1059–1078.
Chambers, C. 
2017.	The Seven Deadly Sins of Psychology: A Manifesto for Reforming the Culture of Scientific Practice. Princeton University Press, Princeton, New Jersey. 
Clarke, M.
2015.	The Digital Dilemma: Preservation and the Digital. Archaeological Record. Advances in Archaeological Practice. 3(4): 313-330.
Cranston K, Harmon L. J., O'Leary, M. A. and C. Lisle
2014.	Best Practices for Data Sharing in Phylogenetic Research. PLOS Currents Tree of Life. (blog) 19 Jun 2014, DOI: 10.1371/currents.tol.bf01eff4a6b60ca4825c69293dc59645, accessed 7 September 2017
Crossref 
2017. Event Data User Guide, Version 1.0.0 - Beta. Electronic document, https://www.eventdata.crossref.org/guide/, accessed 7 November  2017.
Daston, L, and P. Galison
1992.	 The Image of Objectivity. Representations. 40:81–128.
Data Citation Synthesis Group
2014.	Joint Declaration of Data Citation Principles. In FORCE11; 2014 edited by Martone M., Electronic document, https://doi.org/10.25490/a97f-egyk, accessed on 7 September 2017. 
Dorch, B. F., Drachen, T. M., and O. Ellegaard
2015.	The Data Sharing Advantage in Astrophysics. Proceedings of the International Astronomical Union, 11(A29A), 172-175.
Duke, C. S., and J. H. Porter
2013.	The Ethics of Data Sharing and Reuse in Biology. BioScience, 63(6), 483–489. 
Editors
2013.	Guide for Authors. Journal of Archaeological Science. Electronic document, https://web.archive.org/web/20130125222940/https://www.elsevier.com/journals/journal-of-archaeological-science/0305-4403/guide-for-authors, accessed 7 September 7, 2017
Ellison, L. A. 
2012.	Just 20 days left!! Upload Your Archaeological Information FREE to tDAR!! | The Digital Archaeological Record (blog). Electronic document https://www.tdar.org/news/2012/12/just-20-days-left-upload-your-archaeological-information-free-to-tdar/, accessed 7 September 2017
Eglen, S. J., Marwick, B., Halchenko, Y. O., Hanke, M., Sufi, S., Gleeson, P., J.-B. Poline
2017.	Toward Standard Practices for Sharing Computer Code and Programs in Neuroscience. Nature Neuroscience, 20(6), 770–773. 
Fienberg, S. E., Martin, M. E., & Straf, M. L. 
1985	Sharing Research Data. National Academy Press, Washington DC.
Floridi, L. 
2011.	The Philosophy of Information. Oxford University Press, London.
Giofrè, D., Cumming, G., Fresc, L., Boedker, I., and P. Tressoldi
2017.	The influence of journal submission guidelines on authors’ reporting of statistics and use of open research practices. PLoS ONE, 12(4). DOI:10.1371/journal.pone.0175583, accessed on 7 September 2017
Gleditsch, N. P., Metelits, C., and H. Strand
2003.	Posting Your Data: Will You be Scooped or will You be Famous. International Studies Perspectives, 4(1), 89-97.
Hacking I.
1992.	The Self-Vindication of the Laboratory Sciences. In Science as Practice and Culture, edited by Andrew Pickering. The University of Chicago Press, pp. 29–64, Chicago, Illinois. 
Hagstrom, W O. 
1982.	Gift Giving as an Organising Principle in Science In Science in Context: Readings in the Sociology of Science edited by B. Barnes and D. Edge, pp. 21–34, The Open University Press, Milton Keynes. 
Harding, A., Harper, B., Stone, D., O’Neill, C., Berger, P., Harris, S., and J. Donatuto
2012.	Conducting Research with Tribal Communities: Sovereignty, Ethics, and Data-Sharing Issues. Environmental Health Perspectives, 120(1), 6.
Hart, E. M., Barmby, P., LeBauer, D., Michonneau, F., Mount, S., Mulrooney, P., and , J. W. Hollister
2016.	Ten Simple Rules for Digital Data Storage. PLOS Computational Biology, 12(10), e1005097. DOI:10.1371/journal.pcbi.1005097, accessed on 7 September 2017
Henneken, E. A., and A. Accomazzi
2011.	Linking to Data-Effect on Citation Rates in Astronomy. Electronic document, https://arxiv.org/abs/1111.3618, accessed 17 September 2017
Huggett, J. 
2017	Citing Data Reuse. Electronic document, https://introspectivedigitalarchaeology.wordpress.com/2017/05/23/citing-data-reuse/, accessed September 7, 2017
Hymes, D. 
1972.	Reinventing Anthropology. Pantheon, New York.
Inter-university Consortium for Political and Social Research
2012.	Guide to Social Science Data Preparation and Archiving: Best Practice Throughout the Data Life Cycle (5th ed.). Institute for Social Research University of Michigan, Ann Arbor, Michigan
Kansa, E. C. and S. W. Kansa
2013.	Open Archaeology: We All Know That a 14 Is a Sheep: Data Publication and Professionalism in Archaeological Communication. Journal of Eastern Mediterranean Archaeology & Heritage Studies, 1(1), 88-97.
Kansa, E. C., S. W. Kansa and E. Watrall, E. (editors)
2011.	Archaeology 2.0: new approaches to communication and collaboration. Cotsen Digital Archaeology Series, University of California, Electronic document, https://escholarship.org/uc/item/1r6137tb, accessed 7 November 2017
Kidwell, M. C., Lazarević, L. B., Baranski, E., Hardwicke, T. E., Piechowski, S., Falkenberg, L. S. and T. M. Errington
2016.	Badges to Acknowledge Open Practices: A Simple, Low-cost, Effective Method for Increasing Transparency. PLoS Biology, 14(5), e1002456.
Kintigh, K. 
2006.	The Promise and Challenge of Archaeological Data Integration. American Antiquity, 71(3), 567-578.
Keers, G., van der Reijden, H., and H. van Rossum
2011.	Planning Archaeology: A Synthesis of the Thematic Sub-reports. RIGO Research en Advies, Amsterdam. 
Kratz J. and C. Strasser 
2015.	 Researcher Perspectives on Publication and Peer Review of Data. PLOS ONE 10(2): e0117619. DOI:10.1371/journal.pone.0117619, accessed 7 September 2017.
Kratz J. and C. Strasser 
2014.	Data publication consensus and controversies [version 1; referees: 1 approved with reservations]. F1000Research, Electronic document, https://f1000research.com/articles/3-94/v1, accessed 7 September 2017.
Kyzas P.A, K.T. Loizou, and J.P. Ioannidis
2005.	Selective reporting biases in cancer prognostic factor studies. Journal of the National Cancer Institute 97, 1043-1055.
Latour, B. 
1999.	Pandora's Hope: Essays on the Reality of Science Studies. Harvard University Press, Massachusetts. 
Leonelli, S. 
2016.	The Philosophy of Data, In The Routledge Handbook of Philosophy of Information, edited by L. Floridi. Routledge, New York.
Leonelli, S. 
2015.	 What Counts as Scientific Data? A Relational Framework. Philosophy of Science, 82(5), 810-821.
Marwick, B. 
2017.	Computational Reproducibility in Archaeological Research: Basic Principles and a Case Study of their Implementation. Journal of Archaeological Method and Theory, 24(2), 424-450.
Marwick, B., E. Hayes, C. Clarkson and R. Fullagar
2017.	Movement of Lithics by Trampling: An Experiment in the Madjedbebe Sediments, Northern Australia. Journal of Archaeological Science 79:73-85.
McCoy, M. D. and M. C. Codlin
2016.	The Influence of Religious Authority in Everyday Life: A Landscape-Scale Study of Domestic Architecture and Religious Law in Ancient Hawai’i. World Archaeology 48 (3):411–30.
McCullough B. D., K.A. McGeary and T. D. Harrison 
2006.	Lessons from the JMCB Archive. Journal of Money, Credit, and Banking 38(4), 1093-1107.
McCullough B. D. and H. D. Vinod
2005.	Comment: Econometrics and Software. Journal of Economic Perspectives 17(1), 223-224.
McMurry, J. A., Juty, N., Blomberg, N., Burdett, T., Conlin, T., Conte, N., Parkinson, H., Winfree, L., Mungall, C., Haendel, M. and H. Parkinson 
2017.	Identifiers for the 21st Century: How to Design, Provision, and Reuse Persistent Identifiers to Maximize Utility and Impact of Life Science Data. PLOS Biology, 15(6), e2001414. DOI:10.1371/journal.pbio.2001414, accessed on 7 September 2017.
Merton, R.K. 
1942.	The Normative Structure of Science. In The Sociology of Science: Theoretical and Empirical Investigations, edited by R. K. Merton, University Of Chicago Press, Chicago, Illinois. 
Moore, R., and J. Richards
2015.	Here Today, Gone Tomorrow: Open Access, Open Data and Digital Preservation. In Open Source Archaeology: Ethics and Practice, edited by Andrew T. Wilson, Ben Edwards, De Gruyter, Netherlands
Mounce, R. 
2014.	Open Data and Palaeontology. In Issues in Open Research Data, edited by S. A. Moore. pp. 151–164. Ubiquity Press, London.
National Science Foundation
2016.	Dissemination and Sharing of Research Results. Electronic document, https://www.nsf.gov/bfa/dias/policy/dmp.jsp, accessed 7 September 2017
Nicholas, G. P.
2014.	Indigenous cultural heritage in the age of technological reproducibility: Towards a postcolonial ethic of the public domain. In Dynamic Fair Dealing: Creating Canadian Culture Online, edited by Coombe, R. J., Wershler, D., and M. Zeilinger, pp. 213-24. University of Toronto Press, Toronto. 
Nicholas, G. P., and A. Wylie
2012.	Do Not Do Unto Others…’: Cultural misrecognition and the harms of appropriation in an open-source world. In Appropriating the Past: Philosophical Perspectives on Archaeological Practice, edited by R. Coningham and G. Scarre, pp. 195-221, Cambridge University Press, Cambridge.
Nicholas, G. P. and A. Wylie
2009.	Archaeological Finds: Legacies of Appropriation, Modes of Response. The Ethics of Cultural Appropriation, edited by J. O. Young and C. G. Brunk, pp. 11-54, Wiley-Blackwell, Oxford.
Nosek, B. A., Alter, G., Banks, G. C., Borsboom, D., Bowman, S. D., Breckler, S. J. and T.  Yarkoni
2015.	Promoting an Open Research Culture: Author Guidelines for Journals Could Help to Promote Transparency, Openness, and Reproducibility. Science, 348, 1422–5.
Nugent, R., & Williams, H. 
2012.	Sighted surfaces: ocular agency in early Anglo-Saxon cremation burials. In Encountering Imagery: Materialities, Perceptions, Relations, edited by I.-M. Back Danielsson, F. Fahlander and Y. Sjöstrand. pp. 187-208, Stockholm University, Stockholm.
Nuijten, M. B., Borghuis, J., Veldkamp, C. L. S., Alvarez, L. D., van Assen, M. A. L. M., and J. M. Wicherts
2017.	Journal Data Sharing Policies and Statistical Reporting Inconsistencies in Psychology. Electronic document, https://psyarxiv.com/sgbta, accessed 7 September 2017
National Science Foundation
2010.	Data Management & Sharing Frequently Asked Questions (FAQs). Electronic document, https://www.nsf.gov/bfa/dias/policy/dmpfaqs.jsp#1, accessed on 7 September 2017
Pampel, H., Vierkant, P., Scholze, F., Bertelmann, R., Kindling, M., Klump, J. and U. Dierolf
2013	Making research data repositories visible: The re3data.org registry. PloS ONE, 8(11), e78080, DOI: 10.1371/journal.pone.0078080, accessed 7 September 2017.
Pavlech, L. L.
2016	 Data Citation Index. Journal of the Medical Library Association, 104(1), 88–90.
Peng R. D., F. Dominici F, and S. L. Zege
2006.	 Reproducible Epidemiological Research. American Journal of Epidemology 163, 783-789
Penev L, Mietchen D, Chavan V, Hagedorn G, Smith V, Shotton D, Ó Tuama É, Senderov V, Georgiev T, Stoev P, Groom Q, Remsen D, and S. Edmunds
2017	Strategies and guidelines for Scholarly Publishing of Biodiversity Data. Research Ideas and Outcomes 3: e12431. DOI:10.3897/rio.3.e12431, accessed 7 September 2017.
Pienta, A. M., Alter, G. C., and J. A Lyle
2010	The Enduring Value of Social Science Research: The Use and Reuse of Primary Research Data. Electronic document, http://deepblue.lib.umich.edu/handle/2027.42/78307, accessed 27 September 2017
Piwowar H. A., R. S. Day and D. B. Fridsma 
2007.	Sharing Detailed Research Data is Associated with Increased Citation Rate. PLOS ONE 2(3): e308. 
Piwowar, H. A. and T. J. Vision
2013.	Data reuse and the open data citation advantage. PeerJ, 1, e175. DOI:10.7717/peerj.175 accessed on 7 September 2017
PLoS Medicine Editorial and Publishing Policies. 
2017.	Sharing of Materials, Methods and Data. PLoS Medicine Editorial and Publishing Policies. Electronic document. http://journals.plos.org/plosmedicine/policies.php#sharing, accessed on 7 September 2017
Reichman, J. H., and P. F. Uhlir
2003.	A Contractually Reconstructed Research Commons for Scientific Data in a Highly Protectionist Intellectual Property Environment. Law and Contemporary Problems, 66(1/2), 315-462.
Reidpath D. D. and P. A. Allotey
2002.	Data Sharing in Medical Research: An Empirical Investigation. Bioethics 15(2), 125- 134.
Rowhani-Farid, A., Allen, M., and A. G.Barnett
2017.	What Incentives Increase Data Sharing in Health and Medical Research? A Systematic Review. Research Integrity and Peer Review, 2(1), 4. 
Savage C. J. and A. J. Vickers
2009.	Empirical Study of Data Sharing by Authors Publishing in PLoS Journals. PLoS ONE 4(9): e7078. doi:10.1371/journal.pone.0007078, accessed 7 September 2017.
Sawyer, S. 
2008.	Data Wealth, Data Poverty, Science and Cyberinfrastructure. Prometheus, 26(4), 355-371.
Scientific Data Editors
n.d	 Recommended Data Repositories. Electronic document, https://www.nature.com/sdata/policies/repositories, accessed 13 September 2017
Sears, J.R.L 
2011.	Data Sharing Effect on Article Citation Rate in Paleoceanography. Presented at Fall Meeting of the American Geophysical Union. Electronic document, https://doi.org/10.6084/M9.FIGSHARE.1222998.V1, accessed 7 September 2017.
Spencer, H.
2010.	Thoughts on the Sharing of Data and Research Materials and the Role of Journal Policies. Electronic document, https://web.stanford.edu/~vcs/Nov21/hilary_spencer_rdcscsJan2010.pdf, accessed 7 September  2017
Strasser, C., Cook, R., Michener, W. and A. Budden
2012.	Primer on Data Management: What you always wanted to know. DataONE Best Practices Primer. Electronic document, https://doi.org/10.5060/D2251G48, accessed, 7 September 2017 
Stodden, V. 
2010.	The Scientific Method in Practice: Reproducibility in the Computational Sciences, Columbia University Academic Commons, Electronic document, http://hdl.handle.net/10022/AC:P:11417, Accessed 7 September 2017.
Stodden, V. 
2009a.	The Legal Framework for Reproducible Scientific Research: Licensing and Copyright. Computing in Science & Engineering, 11(1), 35-40.
Stodden, V.
2009b.	Enabling Reproducible Research: Open Licensing for Scientific Innovation. International Journal of Communications Law and Policy. https://ssrn.com/abstract=1362040, accessed 7 September 2017.
Tenopir, C., Dalton, E. D., Allard, S., Frame, M., Pjesivac, I., Birch, B and K. Dorsett
2015.	Changes in Data Sharing and Data Reuse Practices and Perceptions among Scientists Worldwide. PLOS ONE, 10(8), e0134826. DOI:10.1371/journal.pone.0134826, accessed 7 September 2017.
Thessen, A. E. and D. J. Patterson 
2011.	Data issues in the life sciences. ZooKeys 150, 15–51. DOI:10.3897/zookeys.150.1766, accessed 7 September 2017. 
Tsosie, R. 
2007. Cultural Challenges to Biotechnology: Native American Genetic Resources and the Concept of Cultural Harm. The Journal of Law, Medicine & Ethics 35(3), 396–411. 
Uhlir, P.F., & Cohen, D. 
2011. Internal document. Board on Research Data and Information, Policy and Global Affairs Division, National Academy of Sciences. 18 March 2011. Washington, DC.
Archaeology Data Service / Digital Antiquity 
2011. Guides to Good Practice. Electronic document, http://guides.archaeologydataservice.ac.uk/ accessed on 7 September 2017. 
Vanpaemel, W., Vermorgen, M., Deriemaecker, L., and G. Storms
2015. 	Are We Wasting a Good Crisis? The Availability of Psychological Research Data after the Storm. Collabra, 1(1): 3. DOI:10.1525/collabra.13, accessed 7 September 2017. 
Vines, T.H, A. Y. Albert, R. L. Andrew, F. Débarre, D. G.Bock, M. T. Franklin and D. J.  Rennison
2014.	 The Availability of Research Data Declines Rapidly with Article Age. Current Biology 24(1):94-97.
Vitelli, K. D. and J. S. Colwell-Chanthaphonh, (editors)
2006.	Archaeological Ethics. Rowman Altamira, New York.
Wallis, J. C., Rolando, E. and C. L. Borgman
2013.	If We Share Data, Will Anyone Use Them? Data Sharing and Reuse in the Long Tail of Science and Technology. PLoS ONE, 8(7), e67332. DOI:10.1371/journal.pone.0067332, accessed 7 September 2017.
White E, Baldridge E, Brym Z, Locey K, McGlinn D and S. Supp
2013.	 Nine simple ways to make it easier to (re)use your data. Ideas in Ecology and Evolution. 2013;6: 1–10. DOI:10.4033/iee.2013.6b.6.f, accessed on 7 September 2017
Whitlock, M. C. 
2011.	Data Archiving in Ecology and Evolution: Best Practices. Trends in Ecology & Evolution, 26(2), 61-65.
Wicherts, J. M., Borsboom, D., Kats, J. and D. Molenaar
2006.	The Poor Availability of Psychological Research Data for Reanalysis. American Psychologist, 61(7), 726.
Wilkinson, M. D., Dumontier, M., Aalbersberg, I. J., Appleton, G., Axton, M., Baak, A. and P. E. Bourne
2016.	The FAIR Guiding Principles for scientific data management and stewardship. Scientific Data, 3, 160018. DOI:10.1038/sdata.2016.18, accessed on 7 September 2017.
Womack R.P.
2015.	Research Data in Core Journals in Biology, Chemistry, Mathematics, and Physics. PLOS ONE 10(12): e0143460. DOI:10.1371/journal.pone.0143460, accessed on 7 September 2017. 
Wouters P. and W. Haak 
2017.	Open Data Report: The Researcher Perspective. Electronic document https://www.elsevier.com/about/open-science/research-data/open-data-report, accessed 11 October 2017
Zimmerman, L. J., Vitelli, K. D., and J. Hollowell-Zimmer, (editors)
2003.	Ethical Issues in Archaeology. Rowman Altamira, New York.

### Colophon

This document contains `r prettyNum(wordcountaddin:::text_stats_fn_(readLines("paper.Rmd"))$n_words_korp, big.mark = ",")` words and was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE, eval = TRUE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r eval = TRUE}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```